---
title: "OkeanosHD-VLC"
output: html_document
---
## File Names to Run Times
This code will create a data frame from your Okeanos HD video file names that has your:  
* File Path
* File Name
* DateTime in UTC (for every second of video)
* RunTime (as 0 to a number like 360 assigned to each second of video, this is needed to grab the screenshot from VLC using the video run time)
```{r}
library(tidyr)
library(stringr)
library(lubridate)
library(dplyr)

# Set working directory to wherever the HD video files are located
setwd("C:/Users/elizabeth.gugliotti/Desktop/ERDAP/Okeanos test")

# Create a vector using that file path for later
filepath<-"C:/Users/elizabeth.gugliotti/Desktop/ERDAP/Okeanos test"

# This creates a vector of your file names in that working directory
files<-list.files()

filex<-unlist(str_split(files,"_",n=5)) # unlist and break string apart by underscore
x<-matrix(filex,ncol=5,byrow=TRUE) # create a matrix
colnames(x)<-c("Dive","Vid","DateTime", "yo","yo1") # add column names
x<-data.frame(x) # make x a data frame

# Break apart the DateTime column in x into the year, month, day, hour, minutes, seconds
year<-substr(x$DateTime ,1,4)
month<-substr(x$DateTime ,5,6)
day<-substr(x$DateTime ,7,8)
hr<-substr(x$DateTime ,10,11)
mints<-substr(x$DateTime ,12,13)
secs<-substr(x$DateTime ,14,15)

# Turn this into a DateTime format
DT<-ISOdatetime(year=year, month=month, day=day, hour=hr, min=mints, sec=secs, tz="UTC")
# you will have to manually insert the end time of your final video. I didn't do that here so i just cut out the last 5 minutes. Honestly it should be when the rov is back on deck so you could probably just omit that last video anyways so you don't have to manually enter it in.

NextDT<-DT[2:3] #create vector that is n+1

#starttime + ((next starttime - starttime) -1) = end time
EndDT<-DT[1:2]+ difftime(NextDT, DT[1:2],units="secs")-1
StartDT<-DT[1:2] #Need vector that is a length of 2 just for this exercise
path<- rep(filepath,length(StartDT)) # Make vector that is the filepath repeated for every row (necessary for VLC conversion)
files<-files[1:2] #Need vector that is length of 2 for this exercise

all<-do.call("cbind",list(path,files, as.character(StartDT),as.character(EndDT))) #use do.call with cbind to combine multiple vectors into one dataframe
colnames(all)<-c("Path","Files","StartDT","EndDT")
all<-data.frame(all)
# You should bind DT (start time), EndDT, your file names, and file paths
# Your file paths should be your working directory which I created a vector called "filepath" and your file names should be the same ones as the "files" vector above

# Even though I put is as.character, it came out as a factor so I need to convert it to a POSIXct class
all$StartDT<-as.POSIXct(all$StartDT)
all$EndDT<-as.POSIXct(all$EndDT)

# Create empty data frame
video.df <- data.frame(Path = character(),
                 File = character(),
                 DateTime = as.POSIXct(character()),
                 RunTime=numeric(),
                 stringsAsFactors=FALSE)

# Create the loop
for (id in all$Files){
  x<- seq.POSIXt(all$StartDT[all$Files==id], all$EndDT[all$Files==id], units="seconds", by= 1)
  f<-as.numeric(difftime(all$EndDT[all$Files==id],all$StartDT[all$Files==id],units="secs"))
  w<- seq(from = 0, to=f, by=1)
  g<- rep(all$Path[all$Files==id],length(x))
  h<- rep(all$Files[all$Files==id],length(x))
  test.d <- data.frame(Path = g,
                  File = h,
                  DateTime = x,
                  RunTime = w,
                  stringsAsFactors=FALSE)
  video.df <- rbind(video.df, video.d)
}
```

## VLC
Using the runtimes you created and your csv annotations we are going to create a .bat file that will run VLC and grab a snapshot at the times of your observations
```{r}
#### Prepping your data ####

# Read in the csv with your annotations
annotations<-read.csv("nameofyourcsv.csv", header = TRUE)

# I have a Date and a Time column so I created onother column with both called DateTime
annotations$DateTime<-paste(annotations$Date,annotations$TC, sep=" ")
annotations$DateTime<-mdy_hms(annotations$DateTime) # Make sure this is actually in a DateTime format that R understands

# Same thing with video data
video$DateTime<-paste(video$Date, video$Time, sep=" ")
video$DateTime<-mdy_hms(video$DateTime)

# Match DateTime in annotations to DateTime in video
time.match<-subset(annotations, (DateTime %in% video$DateTime))
# Match DateTime in video to DateTime in annotations
vtime.match<-subset(video, (DateTime %in% annotations$DateTime))
all.match<-full_join(time.match, vtime.match, by= "DateTime") # Join matches
all.match$RunTime<-as.numeric(all.match$RunTime) # Have to make sure that the RunTime is a numeric vector

# Need to create and EndTime vector so that VLC will know when to stop taking screenshots
all.match$EndTime<-all.match$RunTime + 1



#### Create a file name structure similar to what SeaTube makes ####
#Dates
dates<-unlist(strsplit(as.character(all.match$Date.x), split="/")) # Break apart date
m<-matrix(dates,ncol=3,byrow=TRUE) # Make matrix
colnames(m)<-c("Month","Day","Year") # Name columns
m<-data.frame(m) # Turn matrix to data frame
ymd<-paste(m$Year,m$Month,m$Day, sep="") # Paste it all together so that it is YYYYMMDD

#Times
times<-unlist(strsplit(as.character(all.match$TC), split=":")) # Break apart time
x<-matrix(times,ncol=3,byrow=TRUE) # Make matrix
colnames(x)<-c("Hours","Minutes","Seconds") # Name columns
x<-data.frame(x) # Turn matrix to data frame
hms<-paste(x$Hours,x$Minutes,x$Seconds,sep="") # Paste it all together so that it is HHMMSS

all.match$DTN<-paste(ymd,"T",hms,"_",all.match$ScientificName, sep="") #Paste Date, Time, and ScientificName together for file name

# Make the command
cmd <- paste('vlc',' ', all.match$FilePath, all.match$VideoName,'.mp4',' ',
               '--video-filter=scene --start-time=',all.match$RunTime, ' ',
               '--stop-time=', all.match$EndTime, ' ', '--scene-ratio=24',' ',
               '--scene-prefix=',all.match$DTN,' ',
               '--scene-path=C:\\Users\\elizabeth.gugliotti\\Desktop\\ERDAP\\screenshots', ' ',  # Make sure this is the filepath you want
               'vlc://quit', sep = '')



#### build a *.bat file ####
bat<-data.frame(cmd)

cd<- "cd C:\\Program Files\\VideoLAN\\VLC"
cd<-data.frame(cd)

names(cd)<-names(bat)
bat<-rbind(cd,bat)


write.table(bat, "C:/Users/elizabeth.gugliotti/Desktop/ERDAP/vlc_vidcap.bat",
            quote = F,
            row.names = F,
            col.names = F)

shell.exec("C:/Users/elizabeth.gugliotti/Desktop/ERDAP/vlc_vidcap.bat")
```

